## Tiefe? Masse? Wasse'dasse?



Unterfunktionen:
```{r, help functions}
train_values <- function(data, dimension, numb_rows,
                         subsample, scope, dummy = 1L) {
  if (checkmate::test_data_frame(data)) {
    data <- as.matrix(data)
  }

  direction <- rnorm(dimension)
  subsample_size <- subsample * numb_rows

  sample_rows <- sample(seq_len(numb_rows))[seq_len(subsample_size)]
  sub_data <- data[sample_rows, ]

  projection <- sub_data %*% direction

  maximum <- max(projection)
  minimum <- min(projection)
  midrange <- (maximum + minimum) / 2

  random_selection <- runif(
    1,
    min = midrange - scope * (maximum - minimum) / 2,
    max = midrange + scope * (maximum - minimum) / 2
  )


  number_of_lower_values <- sum(projection < random_selection)
  rate_of_lower_values <- number_of_lower_values / subsample_size

  # It is not necessary to save the rate_of_greater_values as suggested in the paper.
  # Because 1 - rate_of_lower_values = rate_of_greater_values.
  # But 'mass' uses the rate and 'depth' the absolute value,
  # this is why the number_of_lower_values is saved instead.
  list(
    "direction" = direction,
    "random_selection" = random_selection,
    "rate_of_lower_values" = rate_of_lower_values,
    "number_of_lower_values" = number_of_lower_values
  )
}


evaluate_values <- function(halfspace, data, metric = c("mass", "depth")) {
  if (checkmate::test_data_frame(data)) {
    data <- as.matrix(data)
  }
  lower_value <- ifelse(
    metric == "mass",
    halfspace$rate_of_lower_value,
    halfspace$number_of_lower_values
  )

  other_value <- ifelse(
    metric == "mass",
    1 - halfspace$rate_of_lower_values,
    NA
  )

  projection <- data %*% halfspace$direction

  evaluate_value <- ifelse(
    projection < halfspace$random_selection,
    lower_value,
    other_value
  )
  evaluate_value
}
```

Trainingsphase:
```{r,train depth}
train_depth <- function(data,
                        n_halfspace,
                        subsample = 1L,
                        scope = 1L,
                        seed = NULL) {
  checkmate::assert(checkmate::check_data_frame(data), checkmate::check_matrix(data))
  checkmate::assert_count(n_halfspace)
  checkmate::assert_number(subsample, lower = 0, upper = 1)
  checkmate::assert_number(scope, lower = 0)
  checkmate::assert_number(seed, null.ok = TRUE)


  dimension <- ncol(data)

  if (dimension < 2L) {
    stop("Dimension of data must be at least 2.")
  }


  set.seed(seed)

  lapply(
    seq_len(n_halfspace),
    FUN = train_values,
    data = data,
    dimension = dimension,
    numb_rows = nrow(data),
    subsample = subsample,
    scope = scope
  )
}
```

Testphase:
```{r, evaluate depth}
evaluate_depth <- function(data, halfspaces, metric = "mass") {
  checkmate::assert(
    checkmate::check_data_frame(data),
    checkmate::check_matrix(data)
  )
  checkmate::assert_list(halfspaces)
  checkmate::assert_set_equal(
    names(halfspaces[[1]]),
    c(
      "direction", "random_selection",
      "rate_of_lower_values", "number_of_lower_values"
    )
  )

  if (ncol(data) != length(halfspaces[[1]]$direction)) {
    stop("Training and evaluation data are not the same dimension!")
  }

  match.arg(metric, c("mass", "depth"))


  halfspace_mass <- lapply(
    halfspaces,
    FUN = evaluate_values,
    data = data,
    metric = metric
  )
  halfspace_mass <- matrix(unlist(halfspace_mass), ncol = length(halfspace_mass))


  if (metric == "mass") {
    return(rowMeans(halfspace_mass))
  }
  return(apply(halfspace_mass, 1, min, na.rm = TRUE))
}
```


Grafiken aus dem Paper:
```{r, vis-hsm-prep}
library(ggplot2)
theme_set(theme_minimal())

# visualize half-space depth/mass values for 2D-data on a grid of points
# if no grid is provided, a grid over min/max +/- .2 * range of data is created.
#   data: a 2d data.frame with columns z1 and z2
#   points: add points in data to plot?
#   metric: argument for evaluate_depth
plot_depth <- function(halfspaces, data, grid = NULL, points = TRUE,
                       gridlength = 70, metric = "mass") {
  if (is.null(grid)) {
    range_1 <- range(data$z1)
    range_2 <- range(data$z2)
    grid <- expand.grid(
      z1 = seq(range_1[1] - .2 * diff(range_1),
        range_1[2] + .2 * diff(range_1),
        length = gridlength
      ),
      z2 = seq(range_2[1] - .2 * diff(range_2),
        range_2[2] + .2 * diff(range_2),
        length = gridlength
      )
    )
  }
  grid_depth <- evaluate_depth(
    data = as.matrix(grid),
    halfspaces = halfspaces,
    metric = metric
  )
  grid_halfspaces <- cbind(grid, depth = grid_depth)
  # use colors as in Chen et al.:
  spectralcolors <- c(
    "darkblue", "blue", "cyan", "lightgreen",
    "yellow", "orange", "red", "darkred"
  )
  p <- ggplot(grid_halfspaces, aes(x = z1, y = z2)) +
    geom_tile(aes(fill = depth, colour = depth)) +
    scale_fill_gradientn(metric, colors = spectralcolors) +
    scale_colour_gradientn(metric, colors = spectralcolors)

  if (points & !is.null(data)) {
    p <- p +
      geom_point(
        data = data,
        aes(x = z1, y = z2),
        colour = rgb(1, 1, 1, .8)
      )
  }
  p
}
```
Check Figure 3:
```{r, vis-hsm-fig3, fig.width = 8, fig.height = 4, out.width = ".6\\textwidth"}
library(gridExtra)
data_fig3 <- data.frame(z1 = c(-2, -.5, .5, 2), z2 = 0)
grid_fig3 <- expand.grid(
  z1 = seq(-3, 3, l = 51),
  z2 = seq(-3, 3, l = 51)
)

depth_fig3 <- train_depth(data_fig3,
  n_halfspace = 1e4,
  scope = 1, seed = 4163
)
# need scope > 1 for reliable halfspace _depth_ approximation:
depth_fig3_scope15 <- train_depth(data_fig3,
  n_halfspace = 1e3, scope = 1.5,
  seed = 4163
)
gridExtra::grid.arrange(
  plot_depth(depth_fig3_scope15,
    data = data_fig3, grid = grid_fig3,
    metric = "depth"
  ) +
    ggtitle("Tukey Halfspace Depth"),
  plot_depth(depth_fig3, data = data_fig3, grid = grid_fig3) +
    ggtitle("Halfspace Mass (Chen et al.)"),
  nrow = 1
)
# NB: color scale not exactly as in Chen et al, but results seem very close...
```

Check Figure 5: 
```{r, vis-hsm-fig5, fig.width = 8, fig.height = 4, out.width = ".6\\textwidth"}
set.seed(187471431)
# 2D standard Normal:
cluster <- data.frame(
  z1 = rnorm(50) / 2,
  z2 = rnorm(50) / 2,
  group = "cluster"
)
# polar coordinates: points with distance 3 to 5 from the origin, at 90° - 270°:
left_anomalies <- data.frame(
  angle = runif(10, pi / 2, 3 * pi / 2),
  length = runif(10, 3, 5)
)
# convert to cartesian coords
left_anomalies <- with(left_anomalies, data.frame(
  z1 = length * cos(angle),
  z2 = length * sin(angle),
  group = "anomaly"
))
# ~ N_2(\mu = (6,0), \Sigma = I_2)
right_anomalies <- data.frame(
  z1 = rnorm(20) / 5 + 6,
  z2 = rnorm(20) / 5,
  group = "anomaly"
)
data_fig5 <- rbind(
  cluster,
  left_anomalies,
  right_anomalies
)

hs_fig5 <- train_depth(data_fig5[, 1:2],
  n_halfspace = 1e4, subsample = .5,
  seed = 4165
)
fig5 <- plot_depth(hs_fig5, data = data_fig5[, 1:2], points = FALSE)
# can't assign two colour scales to one plot, so plot 2 groups separately:
fig5 +
  geom_point(
    data = subset(data_fig5, group == "cluster"),
    aes(x = z1, y = z2), color = rgb(0, 0, 1, .5)
  ) +
  geom_point(
    data = subset(data_fig5, group == "anomaly"),
    aes(x = z1, y = z2), color = rgb(1, 0, 0, .5)
  )
```

Ein paar Tests:
```{r}
library(testthat)

test_that("Input and Output Checks", {
  test_data_frame <- data.frame(a = rnorm(5), b = rnorm(5))

  # Wrong dimension
  expect_error(train_depth(data.frame(1), 2))

  # Should work
  test_train <- train_depth(test_data_frame, n_halfspace = 10L)

  # train_depth result is of length n_halfpsace
  expect_equal(length(test_train), 10L)

  # Train and test dimension differ
  expect_error(evaluate_depth(data.frame(a = 1, b = 2, c = 3), test_train))

  # Wrong list
  expect_error(evaluate_depth(test_data_frame, list("a" = 2, "g" = 4)))

  # Example for dimension 3
  set.seed(2020)
  train_data3d <- data.frame(
    x = runif(1000),
    y = runif(1000), z = runif(1000)
  )
  test_data3d <- data.frame(
    x = c(0.5, 0.4, 0.7, -1),
    y = c(0.5, 0.4, 0.7, -1), z = c(0.5, 0.4, 0.7, -1)
  )

  train3d <- train_depth(train_data3d, n_halfspace = 2000, seed = 202020)

  mass3d <- evaluate_depth(test_data3d, train3d)
  tukey3d <- evaluate_depth(test_data3d, train3d, metric = "depth")

  expect_true(identical(mass3d, sort(mass3d, decreasing = TRUE)))
  expect_true(identical(tukey3d, sort(tukey3d, decreasing = TRUE)))
})
```
Die Idee beim letzten Test für die dritte Dimension ist, dass für die Trainingsphase
ein 3D Würfel, mit Seitenlänge eins erzeugt wird. 
Der Punkt (1/2, 1/2, 1/2) ist genau die Mitte des Würfels, hier sollte die Tiefe/Masse
am höchsten sein. 
Je weiter weg man von diesen Punkt weg kommt desto niedriger ist die Tiefe/Masse,
bis außerhalb des Würfels.
